---
title: "Linguistic Data"
author: "Imen, Khushleen, Abigail"
date: "2023-02-21"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data Introduction
*Dataset exploring state examinations in netherlands as dutch as a second language*
Reference: Schepens, Job, Roeland van Hout, and T. Florian Jaeger. “Big Data Suggest Strong Constraints of Linguistic Similarity on Adult Language Learning.” Cognition 194 (January 1, 2020): 104056. https://doi.org/10.1016/j.cognition.2019.104056.
Data: 
https://www.kaggle.com/datasets/thedevastator/adult-language-learning-profile

Column Name Explanations:  
  
L1: The first language of the learner  
C: The country of birth  
L1L2: The combination of first and best additional language besides Dutch  
AaA: Age at Arrival in the Netherlands in years (starting date of residence)  
LoR: Length of residence in the Netherlands in years  
Edu.day: Duration of daily education (1 low, 2 middle, 3 high, 4 very high)  
Sex: Gender of the participant  
Family: Language family  
ISO639.3: Language ID code according to Ethnologue  
Enroll: Proportion of school-aged youth enrolled in secondary education according to the World Bank for participants country  
Speaking: Speaking proficiency test score on the State Examination of Dutch as a Second Language  
morph: Morphological score related to knowledge structures within words  
lex: Lexicon score indicating understanding of written words  
new_feat: Feature score reflecting ability to acquire new sounds/grammatical structures  
new_sounds: Sound symbols score evaluating pronunciation. (Integer)  

## Who did what?
Imen: Q0 - Data cleaning, Q1-4, 13,14   
Khushleen: Q 5,6,11,12 (and sub-questions)  
Abigail:  Q 7-10  

## 0) Cleaning the data
```{r}
#Loading the packages
library(tidyverse)
library(tidyr)
library(janitor)
library(naniar)
library(dplyr) 
library(ggthemes)
```

```{r}
#install.packages("ggthemes")
```


```{r}
#Load the data
languages <- readr::read_csv("data/stex.csv") 
```
```{r}
# Cleaning Data
languages2<- languages %>% select(-"L1L2") %>% 
  plyr::rename(c("C"="Country", "lex"="Lexicon", "morph"="Morphology","new_feat"="New_Features", "new_sounds"="New_Sounds","Edu.day"="Edu_Days"))
```
 
```{r}
languages2
```

## 1) Where the NAs are:
```{r}
languages2 %>% 
  summarise_all(~(sum(is.na(.))))
```

## 2) Distribution of Sex of Participants
```{r}
languages2 %>% 
  count(Sex) %>% 
  ggplot(aes(x=Sex,y=n, fill=Sex))+
  geom_col()+
  labs(x="Sex", y="Number of Participants", title = "Representation in the Study by Sex")+
  theme_clean()
```

## 3) Distribution of Second Language
```{r}
languages2 %>% 
  count(L2) %>% 
  ggplot(aes(x=L2,y=n, fill=L2))+
  geom_col()+
  labs(x="L2", y="count", title = "Representation of Second Language in the Study Language")+
  scale_y_log10()+
  coord_flip()+
  guides(fill="none")+
  theme_clean()
```

```{r}
languages2 %>% 
  mutate(AaA=as.factor(AaA))
```

## 4) Distribution of Age at Arrival
```{r}
languages2 %>% 
  count(AaA) %>% 
  ggplot(aes(x=AaA,y=n))+
  geom_point()+
  labs(x="Age at Arrival", y="Number of Participants", title = "Representation of Age at Arrival in the Netherlands")+
  scale_y_log10()+
  theme_clean()
```
```{r}
languages2 %>% 
  group_by(AaA) %>% 
  summarise(mean_age_at_arrival=mean(AaA))
```

# 5) Comparing speaking proficiency by age at arrival

- There is a general downward trend; as age at arrival increases, proficiency decreases
```{r}
languages2 %>% 
  group_by(AaA) %>% 
  summarise(Avg_Speaking = mean(Speaking)) %>% 
  ggplot(aes(x=AaA, y=Avg_Speaking))+
  geom_point(shape=2)+
  geom_smooth(method = lm)+
  labs(title = "Speaking Scores By Participants Age at Arrival in Netherlands", x = "Age of Participants", y="Average Speaking Score" )+
  guides(fill="none")+
  theme_clean()
```

# 6) Comparing Length of Education and Speaking Proficiency
```{r}
languages2 %>% 
  mutate(Edu_Days_Category = case_when(Edu_Days == 1 ~ "low",
                   Edu_Days == 2 ~ "middle",
                   Edu_Days == 3 ~ "high",
                   Edu_Days == 4 ~ "very high")) %>% 
  filter(!is.na(Edu_Days_Category)) %>% 
  ggplot(aes(x=Speaking, fill=Edu_Days_Category))+
  geom_histogram(bins = 40, alpha=0.6)+
  facet_wrap(.~Edu_Days_Category)+
  labs(title = "Speaking Scores By Days of Formal Education", x = "Speaking Score", y="Number of Participants")+
  guides(fill="none")+
  theme_clean()
```

- By gender
```{r}
languages2 %>% 
  mutate(Edu_Days_Category = case_when(Edu_Days == 1 ~ "low",
                   Edu_Days == 2 ~ "middle",
                   Edu_Days == 3 ~ "high",
                   Edu_Days == 4 ~ "very high")) %>% 
  filter(!is.na(Edu_Days_Category)) %>% 
  ggplot(aes(x=Speaking, fill=Sex))+
  geom_histogram(bins = 40, alpha = 0.6)+
  facet_grid(Sex~Edu_Days_Category, scales = "free_y")+
  labs(title = "Speaking Scores By Days of Formal Education and Sex", y = "Speaking Score" , x=NULL)+
  guides(fill="none")+
  theme_clean()
```

## 7) Exploring Language Distribution (L1) w/ Country

```{r}
languages2 %>% 
  group_by(Country) %>% 
  select(L1) %>% 
  summarize(n_languages=n_distinct(L1)) %>%  
  arrange(-n_languages) #for top 5 countries that speak the most amount of different L1
```

```{r}
netherlands_languages <- languages2 %>% 
  select(L1, L2, Country) %>% 
  filter(Country == "Netherlands")
netherlands_languages
```

```{r}
p <- netherlands_languages %>% 
  select(L1) %>% 
  group_by(L1) %>% 
  summarize(n=n()) %>% 
  arrange(-n) %>% 
  head(n=10)
p
```

```{r}
p %>% 
  ggplot(aes(x=reorder(L1, n), y=n, fill=L1))+
  geom_col(na.rm = T, alpha=0.6)+ 
  labs(title = "Most Common First Languages (L1) of People Born in the Netherlands",
       x = "Language",
       y = "Count of Residents That Speak L1") +
  theme_clean()+
  guides(fill="none")
```


## 8) Top 10 L2 in Netherlands

```{r}
netherlands_languages %>% 
  select(L2) %>% 
  group_by(L2) %>% 
  summarize(n=n()) %>% 
  arrange(-n) #when exploring top 10 L2, find that 223 residents are monolingual
```

```{r}
p2 <- netherlands_languages %>% 
  select(L2) %>% 
  filter(L2!="Monolingual") %>% 
  group_by(L2) %>% 
  summarize(n_L2=n()) %>% 
  arrange(-n_L2) %>% 
  head(n=10)
p2
```

```{r}
netherlands_languages %>% 
  group_by(Country) %>% 
  summarize(across(c(L1, L2), n_distinct))
```

```{r}
p2 %>% 
  ggplot(aes(x=reorder(L2, n_L2), y=n_L2, fill=L2))+
  geom_col(na.rm = T, alpha=0.6)+ 
  labs(title = "10 Best Additional Languages (L2) of People Born in the Netherlands",
       x = "Language",
       y = "Count of Residents That Speak L2") +
  theme_clean()+
  guides(fill="none")
```

## 9) Morphology & L1 vs Proficiency

```{r}
p4 <- languages2 %>% 
  group_by(Country) %>% 
  filter(Country=="USSR") %>% 
  select(Country, L1, Speaking, Morphology) %>% 
  filter(!is.na(Morphology)) %>% 
  filter(Speaking>="600") %>% 
  arrange(-Speaking)
```

```{r}
p4 %>% 
  ggplot(aes(x=Morphology, y=Speaking))+
  geom_point()+
  facet_grid(.~L1)+
  labs(title = "USSR Proficiency Score vs. Morphological Similarities & L1",
       x= "Score of Morphological Simiarities",
       y= "Proficiency Score")+
  theme_clean()+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## 10) Speaking prof vs. Country & Education Time?
Based off of the top 3 countries that speak the most different amt of L1
Extremes of edu.days (low & very high)

```{r}
p3 <- languages2 %>% 
  group_by(Country) %>% 
  filter(Country=="Netherlands" | 
         Country=="USSR" |
         Country=="Germany") %>% 
  select(Country, Speaking, Edu_Days) %>% 
  mutate(Edu_Days_Category = case_when(Edu_Days == 1 ~ "low",
                   Edu_Days == 2 ~ "middle",
                   Edu_Days == 3 ~ "high",
                   Edu_Days == 4 ~ "very high")) %>% 
  filter(Edu_Days_Category=="very high" |
           Edu_Days_Category=="low")
p3
```

```{r}
p3 %>% 
  ggplot(aes(x=Speaking))+
  geom_density()+
  facet_grid(Country~Edu_Days_Category)+
  theme_clean()+
  labs(title = "Proficiency Score vs. Country & Daily Education Duration",
       x = "Proficiency Score")
```

## 11) LoR vs Speaking Proficiency

- Looking at number of participants per residency length category
```{r}
languages2 %>% 
  mutate(LoR_Category = case_when(LoR <= 15 ~ "low",
                  LoR > 15 & LoR <= 30 ~ "middle",
                  LoR > 30 & LoR <= 45 ~ "high",
                  LoR > 45 ~ "very high")) %>% 
  count(LoR_Category) %>% 
  ggplot(aes(x=LoR_Category, y=n, fill=LoR_Category))+
  geom_col(alpha = 0.6)+
  labs(title = "Participants Per Residency Category", x= "Length of Residency", y = "Number of Participants")+
  guides(fill="none")+
  theme_clean()
```

- Looking at spread of speaking score for LoR category with most participants (low)
```{r}
languages2 %>% 
  mutate(LoR_Category = case_when(LoR <= 15 ~ "low",
                  LoR > 15 & LoR <= 30 ~ "middle",
                  LoR > 30 & LoR <= 45 ~ "high",
                  LoR > 45 ~ "very high")) %>% 
  filter(LoR_Category == "low") %>% 
  ggplot(aes(x=Speaking))+
  geom_histogram()+
  labs(title = "Speaking Score Distribution for Low Residency Category", x= "Speaking Score", y = "Number of Participants")+
  guides(fill="none")+
  theme_clean()
```

- Looking at spread of the other residency categories
```{r}
languages2 %>% 
  mutate(LoR_Category = case_when(LoR <= 15 ~ "low",
                  LoR > 15 & LoR <= 30 ~ "middle",
                  LoR > 30 & LoR <= 45 ~ "high",
                  LoR > 45 ~ "very high")) %>% 
  filter(LoR_Category != "low") %>% 
  ggplot(aes(x=Speaking, fill=LoR_Category))+
  geom_histogram(alpha=0.6, bins=60)+
  facet_wrap(.~LoR_Category)+
  labs(title = "Speaking Scores per Residency Category", x= "Speaking Score", y = "Number of Participants")+
  guides(fill="none")+
  theme_clean()
```

## 12) Comparision of new sounds learned vs language family

- average new sounds per language family
```{r}
languages2 %>% 
  filter(!is.na(New_Sounds)) %>% 
  group_by(Family) %>% 
  summarise(avg_new_sounds = mean(New_Sounds, na.rm = T)) %>% 
  ggplot(aes(x=reorder(Family, -avg_new_sounds) , y=avg_new_sounds, fill=Family))+
  geom_col(alpha = 0.6)+
  labs(title = "Average New Sounds Learned By Language Family", x= "Language Family", y = "Average Number Of New Sounds")+
  guides(fill="none")+
  theme_clean()+
  theme(axis.text.x = element_text(angle = 25, hjust = 0.6))
```

- exploring speaking proficiency scores by language family for least and most new sounds, using relative frequency of scores
```{r}
languages2 %>% 
  filter(Family == "Kartvelian" | Family == "Uralic") %>% 
  ggplot(aes(x=Speaking))+
  geom_histogram(bins = 40, aes(y=after_stat(density), fill=Family), alpha=0.6)+
  facet_wrap(.~Family)+
  labs(title = "Speaking Scores for Family With Least And Most New Sounds", y = "Relative Frequency Of New Sounds")+
  guides(fill="none")+
  theme_clean()
```

## 13) From what countries do people learn Dutch the easiest? (What country has the highest lexicon?)
```{r}
#Highest Lexile(best understanding of new written words on avg):
languages2 %>% 
  group_by(Country) %>% 
  summarise(mean_lexile=mean(Lexicon), na.rm=T) %>% 
  arrange(desc(mean_lexile)) %>% slice_head()
#Highest Morphology(best knowledge of word structure on avg):
languages2 %>% 
  group_by(Country) %>% 
  summarise(mean_morph=mean(Morphology), na.rm=T) %>% 
  arrange(desc(mean_morph)) %>% slice_head() 
#Highest Proficiency(best speaking scores on avg):
languages2 %>% 
  group_by(Country) %>% 
  summarise(mean_speaking=mean(Speaking)) %>% 
  arrange(desc(mean_speaking))%>% slice_head()
```

## 14) Do men or women tend to have an easier time learning Dutch on average?
```{r}
languages2 %>%
  filter(Sex=="Female") %>% 
  summarise(mean_speaking=mean(Speaking), mean_morph=mean(Morphology, na.rm=T), mean_lexile=mean(Lexicon), mean_features=mean(New_Features, na.rm = T))
  
```
```{r}
languages2 %>%
  filter(Sex=="Male") %>% 
  summarise(mean_speaking=mean(Speaking), mean_morph=mean(Morphology, na.rm=T), mean_lexile=mean(Lexicon), mean_features=mean(New_Features, na.rm = T))
```



