---
title: "Linguistic Data"
author: "Imen, Khushleen, Abigail"
date: "2023-02-21"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Column Name Explanations
L1: Native language of the participant. (String)
C: Country of origin of the participant. (String)
L1L2: Linguistic similarity between the native language and the target language. (Integer)
L1L2: Linguistic similarity between the native language and the target language. (Integer)
AaA: Age at arrival in the Netherlands. (Integer)
LoR: Length of residence in the Netherlands. (Integer)
Edu.day: Formal education days in the target language. (Integer)
Sex: Gender of the participant. (String)
Family: Family status of the participant. (String)
ISO639.3: ISO 639-3 codes for the target language. (String)
Enroll: Duration enrolled in language courses. (Integer)
Speaking: Speaking proficiency test score on the State Examination of Dutch as a Second Language. (Integer)
morph: Morphological score related to knowledge structures within words. (Integer)
lex: Lexicon score indicating understanding of written words. (Integer)
new_feat: Feature score reflecting ability to acquire new sounds/grammatical structures. (Integer)
new_sounds: Sound symbols score evaluating pronunciation. (Integer)

```{r}
library(tidyverse)
library(tidyr)
library(janitor)
library(naniar)
library(dplyr) 
```

```{r}
languages <- readr::read_csv("stex.csv") 
```
```{r}
languages2<- languages %>% select(-"L1L2") %>% 
  plyr::rename(c("C"="Country", "lex"="Lexicon", "morph"="Morphology","new_feat"="New_Features", "new_sounds"="New_Sounds","Edu.day"="Edu_Days"))
```
 
```{r}
languages2
```

##Where the NAs are:
```{r}
languages2 %>% 
  summarise_all(~(sum(is.na(.))))
```
##Distribution of Sex
```{r}
#languages2 %>% count("Sex") %>% 
#  ggplot(aes(x=Sex,y=freq, fill=Sex))+geom_col()+labs(x="Sex", y="Number of Participants", title = "Representation in the Study by Sex")+geom_text(aes(label=freq),vjust = -0.5)
```

##Distribution of Second Language
```{r}
#languages2 %>% count("L2") %>% 
 # ggplot(aes(x=L2,y=freq, fill=L2))+geom_col()+labs(x="L2", y="count", title = "Representation of Second Language in the Study Language")+scale_y_log10()+coord_flip()+geom_text(aes(label=freq))
```

```{r}
languages2 %>% 
  mutate(AaA=as.factor(AaA))
```
##Distribution of Age at Arrival
```{r}
#languages2 %>% count("AaA") %>% 
#  ggplot(aes(x=AaA,y=freq))+geom_point()+labs(x="Age at Arrival", y="Number of Participants", title = "Representation of Age at Arrival in the Netherlands")+scale_y_log10()
```
```{r}
languages2 %>% 
  group_by(AaA) %>% 
  summarise(mean_age_at_arrival=mean(AaA))
```


