---
title: "Linguistic Data"
author: "Imen, Khushleen, Abigail"
date: "2023-02-21"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Column Name Explanations
L1: Native language of the participant. (String)
C: Country of origin of the participant. (String)
L1L2: Linguistic similarity between the native language and the target language. (Integer)
L1L2: Linguistic similarity between the native language and the target language. (Integer)
AaA: Age at arrival in the Netherlands. (Integer)
LoR: Length of residence in the Netherlands. (Integer)
Edu.day: Formal education days in the target language. (Integer)
Sex: Gender of the participant. (String)
Family: Family status of the participant. (String)
ISO639.3: ISO 639-3 codes for the target language. (String)
Enroll: Duration enrolled in language courses. (Integer)
Speaking: Speaking proficiency test score on the State Examination of Dutch as a Second Language. (Integer)
morph: Morphological score related to knowledge structures within words. (Integer)
lex: Lexicon score indicating understanding of written words. (Integer)
new_feat: Feature score reflecting ability to acquire new sounds/grammatical structures. (Integer)
new_sounds: Sound symbols score evaluating pronunciation. (Integer)

```{r}
library("tidyverse")
library("tidyr")
```

```{r}
languages <- readr::read_csv("stex.csv")
languages
```
 
```{r}
#languages<-languages %>% 
  #select(-"L1L2")
#languages
```

