---
title: "Linguistic Data"
author: "Imen, Khushleen, Abigail"
date: "2023-02-21"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data Introduction
*Dataset exploring state examinations in netherlands as dutch as a second language*
Reference: Schepens, Job, Roeland van Hout, and T. Florian Jaeger. “Big Data Suggest Strong Constraints of Linguistic Similarity on Adult Language Learning.” Cognition 194 (January 1, 2020): 104056. https://doi.org/10.1016/j.cognition.2019.104056.
Data: 
https://www.kaggle.com/datasets/thedevastator/adult-language-learning-profile

##Column Name Explanations
L1: The first language of the learner
C: The country of birth
L1L2: The combination of first and best additional language besides Dutch
AaA: Age at Arrival in the Netherlands in years (starting date of residence)
LoR: Length of residence in the Netherlands in years
Edu.day: Duration of daily education (1 low, 2 middle, 3 high, 4 very high)
Sex: Gender of the participant
Family: Language family
ISO639.3: Language ID code according to Ethnologue
Enroll: Proportion of school-aged youth enrolled in secondary education according to the World Bank for participants country
Speaking: Speaking proficiency test score on the State Examination of Dutch as a Second Language
morph: Morphological score related to knowledge structures within words
lex: Lexicon score indicating understanding of written words
new_feat: Feature score reflecting ability to acquire new sounds/grammatical structures. (Integer)
new_sounds: Sound symbols score evaluating pronunciation. (Integer)

```{r}
library(tidyverse)
library(tidyr)
library(janitor)
library(naniar)
library(dplyr) 
library(ggthemes)
```

```{r}
#install.packages("ggthemes")
```


```{r}
languages <- readr::read_csv("stex.csv") 
```
```{r}
languages2<- languages %>% select(-"L1L2") %>% 
  plyr::rename(c("C"="Country", "lex"="Lexicon", "morph"="Morphology","new_feat"="New_Features", "new_sounds"="New_Sounds","Edu.day"="Edu_Days"))
```
 
```{r}
languages2
```

##Where the NAs are:
```{r}
languages2 %>% 
  summarise_all(~(sum(is.na(.))))
```

##Distribution of Sex of Participants
```{r}
languages2 %>% 
  count(Sex) %>% 
  ggplot(aes(x=Sex,y=n, fill=Sex))+
  geom_col()+
  labs(x="Sex", y="Number of Participants", title = "Representation in the Study by Sex")#+
  #geom_text(aes(label=freq),vjust = -0.5)
```

##Distribution of Second Language
```{r}
languages2 %>% 
  count(L2) %>% 
  ggplot(aes(x=L2,y=n, fill=L2))+
  geom_col()+
  labs(x="L2", y="count", title = "Representation of Second Language in the Study Language")+
  scale_y_log10()+
  coord_flip()+
  guides(fill="none")#+
  #geom_text(aes(label=freq))
```

```{r}
languages2 %>% 
  mutate(AaA=as.factor(AaA))
```

##Distribution of Age at Arrival
```{r}
languages2 %>% 
  count(AaA) %>% 
  ggplot(aes(x=AaA,y=n))+
  geom_point()+
  labs(x="Age at Arrival", y="Number of Participants", title = "Representation of Age at Arrival in the Netherlands")+
  scale_y_log10()
```
```{r}
languages2 %>% 
  group_by(AaA) %>% 
  summarise(mean_age_at_arrival=mean(AaA))
```
#Comparing speaking proficiency by age
- Looking at the distribution
```{r}
languages2 %>% 
  ggplot(aes(x=Sex, y=Speaking, fill=Sex))+
  geom_boxplot(alpha=0.6)+
  coord_flip()+
  labs(title = "Range of Speaking Scores By Sex of Participants", x = "Speaking Score", y="Sex of Participant" )+
  guides(fill="none")
```
```{r}
languages2 %>% 
  ggplot(aes(x=Speaking, fill=Sex))+
  geom_histogram(alpha=0.6)+
  facet_wrap(.~Sex)+
  labs(title = "Speaking Scores By Sex of Participants", x = "Speaking Score", y="Sex of Participant" )+
  guides(fill="none")
```
#Comparing Length of Education and Speaking Proficiency
```{r}
languages2 %>% 
  mutate(Edu_Days_Category = case_when(Edu_Days == 1 ~ "low",
                   Edu_Days == 2 ~ "middle",
                   Edu_Days == 3 ~ "high",
                   Edu_Days == 4 ~ "very high")) %>% 
  filter(!is.na(Edu_Days_Category)) %>% 
  ggplot(aes(x=Speaking, fill=Edu_Days_Category))+
  geom_histogram(bins = 40, alpha=0.6)+
  facet_wrap(.~Edu_Days_Category)+
  labs(title = "Speaking Scores By Days of Formal Education", y = "Speaking Score", x=NULL )+
  guides(fill="none")
```
- By gender
```{r}
languages2 %>% 
  mutate(Edu_Days_Category = case_when(Edu_Days == 1 ~ "low",
                   Edu_Days == 2 ~ "middle",
                   Edu_Days == 3 ~ "high",
                   Edu_Days == 4 ~ "very high")) %>% 
  filter(!is.na(Edu_Days_Category)) %>% 
  ggplot(aes(x=Speaking, fill=Sex))+
  geom_histogram(bins = 40, alpha = 0.6)+
  facet_grid(Sex~Edu_Days_Category, scales = "free_y")+
  labs(title = "Speaking Scores By Days of Formal Education and Sex", y = "Speaking Score" , x=NULL)+
  guides(fill="none")
```

#Exploring Language Distribution (L1) w/ Country

```{r}
languages2 %>% 
  group_by(Country) %>% 
  select(L1) %>% 
  summarize(n_languages=n_distinct(L1)) %>%  
  arrange(-n_languages) #for top 5 countries that speak the most amount of different L1
```

```{r}
netherlands_languages <- languages2 %>% 
  select(L1, L2, Country) %>% 
  filter(Country == "Netherlands")
netherlands_languages
```

```{r}
p <- netherlands_languages %>% 
  select(L1) %>% 
  group_by(L1) %>% 
  summarize(n=n()) %>% 
  arrange(-n) %>% 
  head(n=10)
p
```
```{r}
netherlands_languages %>% 
  summarise(n_netherland_residents = n()) 
```


```{r}
p %>% 
  ggplot(aes(x=reorder(L1, n), y=n))+
  geom_col(na.rm = T)+ 
  labs(title = "10 Most Common First Languages (L1) in the Netherlands",
       x = "Language",
       y = "Number of Residents That Speak L1") +
  theme_clean()
```

#Top 10 Second Languages in Netherlands

```{r}
netherlands_languages %>% 
  select(L2) %>% 
  group_by(L2) %>% 
  summarize(n=n()) %>% 
  arrange(-n) %>% 
  head(n=10) #when exploring top 10 L2, find that 223 residents are monolingual
```

```{r}
p2 <- netherlands_languages %>% 
  select(L2) %>%  
  filter(L2!="Monolingual") %>% 
  group_by(L2) %>% 
  summarize(n_L2=n()) %>% 
  arrange(-n_L2) %>% 
  head(n=10) #top 10 L2 when "monolingual" is removed
p2
```


```{r}
netherlands_languages %>% 
  group_by(Country) %>% 
  summarize(across(c(L1, L2), n_distinct))
```


```{r}
p2 %>% 
  ggplot(aes(x=reorder(L2, n_L2), y=n_L2))+
  geom_col(na.rm = T)+ 
  labs(title = "10 Best Additional Languages (L2) in the Netherlands",
       x = "Language",
       y = "Number of Residents That Speak L2") +
  theme_clean()
```

#Correlation between L1 & L2 (Top 10 of Each)? Given L1, is there a Predictable L2?



#Speaking prof vs. Country & Education Time?
Based off of the top 3 countries that speak the most different amt of L1
Extremes of edu.days (low & very high)

```{r}
p3 <- languages2 %>% 
  group_by(Country) %>% 
  filter(Country=="Netherlands" | 
         Country=="USSR" |
         Country=="Germany") %>% 
  select(Country, Speaking, Edu_Days) %>% 
  mutate(Edu_Days_Category = case_when(Edu_Days == 1 ~ "low",
                   Edu_Days == 2 ~ "middle",
                   Edu_Days == 3 ~ "high",
                   Edu_Days == 4 ~ "very high")) %>% 
  filter(Edu_Days_Category=="very high" |
           Edu_Days_Category=="low")
p3
```

```{r}
p3 %>% 
  ggplot(aes(x=Speaking))+
  geom_density()+
  facet_grid(Country~Edu_Days_Category)+
  theme_clean()+
  labs(title = "Proficiency Score vs. Country & Daily Education Duration",
       x = "Proficiency Score")
```

